\textcolor[rgb]{0,0,1}{Data-Driven Model Predictive Control using Random Forests for Building Energy Optimization and Climate Control}

### 2018-01-22 paper/final/dpc.tex
reviewed Intro, Sec 2,3,4,5, conclusion and appendix

\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning on historical datasets of buildings and formulate the optimal control problem in a way that Receding Horizon Control (RHC) can be solved efficiently. Let an historical dataset $(\X,\Y)$ be given.
We define $\X = \{(x(k),u(k),d(k))\}_{k=1}^n$ the set of predictor variables (or features), i.e. the set of samples $(x(k),u(k),d(k))$ measured a time instants $k=1,\ldots,n$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the \emph{measurable} state variables (e.g. the rooms temperatures and power consumption of the building), $u(k)\in\mathbb{R}^{n_u}$ is the vector of the \emph{measurable} input variables (e.g. the set-points and schedule information of the building) and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the \emph{measurable and predictable} disturbance variables (e.g. weather historical data).
We define $\Y = \{y(k)\}_{k=1}^{n}$ the set of response variables, i.e. the set of measured samples $y(k)\in\mathbb{R}^{n_y}$ representing the variables we wish to predict with our model. In the most general case, when we want to predict all measured variables, the response variables are represented by the state variables at the next time step, i.e. $y(k) = x(k+1)$. However, in general, we only wish to predict a subset $y(k) = \bar x(k+1) \subset \mathbb{R}^{n_x}$ of variables, e.g. only room temperatures and power consumption.
Clearly, $|\X| = |\Y| = n$ is the number of time samples of the dataset.\\
We remark that, as explained in Section \ref{secModelingIssuesExample}, the dataset $(\X,\Y)$ does not contain non-measurable states such as temperature of wall layers, ceiling, floor, etc.: nevertheless, our simulated and experimental results show that our data-driven methodology is able to compensate the absence of such variables while providing excellent prediction accuracy.}

@@

\textcolor[rgb]{0,0,1}{Assume that, at time $k$, we want to predict the response variable at time $k+j$, i.e. $y(k+j),\ j\in \{1,\ldots,N\}$: when the data has lots of features interacting in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.}

@@

\textcolor[rgb]{0,0,1}{Our modification of the Regression Tree algorithm first partitions the features set $\X$ into the sets $\X^c = \{u(k)\}_{k=1}^n\subset\X$, containing the control (or manipulated) variables, and $\X^d = \{(x(k),d(k))\}_{k=1}^n\subset\X$, containing the disturbance and state (or non-manipulated) variables. The  union of the two non-intersecting sets forms the full feature set of training $\X \equiv \X^c \cup \X^d$.
%Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
%These measurements could directly represent one or more states in the model-based control framework. We denote these as responses $\Y \in \R$ for training, i.e. a $\Y$ represents a particular response and we can have separate models for multiple responses. We define the number of training samples by $|(\X,\Y)| = n$.
Then the training process is divided into the following two steps, which generate as output a tree $\mathcal{T}_j$ as illustrated in Figure~\ref{F:dpc-sepvars} (left):
\begin{enumerate}
	\item the splitting of the dataset only partitions $\X^d$ (see Appendix 1 and \cite{hastie2009elements} for technical details): this choice is necessary to make our model suitable for control, as we will clarify later on, and also reduces the computational complexity. To each leaf $\ell_i$ will correspond an equivalence class of data samples $\X^d_i$ of the partition of $\X^d$, with $\X^d_i\subset \X^d$: as a consequence at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$
	\begin{equation}\label{E:model_tree}
		\ell_i = \mathit{f}_{\mathcal{T}_j} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)
	\end{equation}
	if and only if $( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)) \in \X^d_i$.
	\item in each leaf $\ell_i$ of the tree $\mathcal{T}_j$ we derive, solving a convex program over the data samples in the partition $\X^d_i$ associated to leaf $\ell_i$, an affine function, with coefficients $\beta_{ij}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
	\begin{equation}\label{E:model_leaf}
		y(k+j) =  \beta_{ij} [1,u(k),\ldots,u(k+j) ]^T
	\end{equation}
	Clearly the coefficients $\beta_{ij}$ are different for each leaf $\ell_i$ as they are derived from different sets of samples.
\end{enumerate}
Applying the above procedure for $j=0,\ldots,N$ we build $N$ regression trees $\mathcal{T}_0,\ldots,\mathcal{T}_N$: thus, we have managed to linearize the original model dynamics via black-box modeling. Our two-step training procedure, described by the pseudo code of lines 1-11 of Algorithm~\ref{A:dpcrt}, can be computed off-line: this is an important advantage because the time required to create the model does not affect the control execution in run-time.}

@@

\begin{problem}\label{P:dpcrt}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}     =   \beta_{ij} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                   \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										   \\
		&                                       & & \epsilon_j \geq  0							                                       \\
		&                                       & & j           =    0,\ldots,N.            									       \\
		\end{aligned}
		\label{E:dpcrt}
	\end{equation}
\end{problem}

\textcolor[rgb]{0,0,1}{Here, $Q \succeq 0 \in \mathbb{R}^{n_y\times n_y}$ and $R \succeq 0 \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to tradeoff the importance we provide in the minimization of $y$ versus $u$. The slack variables $\epsilon_j$ are added to ensure recursive feasibility: we relax the equality constraint on $y$ allowing violation up to $\epsilon_j$ to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k \geq 0$. The weight $\lambda$ is then used to tradeoff the importance we provide in the optimal solution to the constraints violation versus the quadratic term. For example in the following sections we will use $\epsilon_j$ and $\lambda$ to define our tolerance to the violation of prescribed bounds by the rooms' temperature.\\
Clearly different cost functions can be chosen depending upon the application, i.e. they can be linear, nonlinear, etc., obviously changing the complexity of the problem. In the current formulation, the data-driven control problem, is reduced to a convex program which is very easy and efficient to solve in C, C++, Python and Matlab, and thus is easily integrable in SCADA systems. Indeed Problem \ref{P:dpcrt} is solved as in the classical MPC formulation, i.e. at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied to the system: $u(k) = u^*_k$.\\
Note that each tree $\T_j$ contributes to Problem  \ref{P:dpcrt} with the linear constraint $y_{k+j} = \beta_{ij} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top$ as a replacement for the state dynamics in the classical MPC formulation. As a consequence, when solving Problem \ref{P:dpcrt} at time $k$, we need to determine the affine functions $\beta_{ij}$: it follows by Equation \eqref{E:model_tree} that, to determine $\beta_{ij}$ at time $k$, the knowledge of the state and disturbance measurements $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) \right)$ is needed, however, values of $d(k+j-\delta_d),\ldots,d(k+j),\ j=0,\ldots,N$ are not available at time $k$, so we need to use disturbance forecast $\tilde d(k+N),\ldots,\tilde d(k)$. Using this information we can narrow down to a leaf in each tree using \eqref{E:model_tree} and thus retrieve the affine model with $\beta_{ij}$ in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon. The run-time solution of Problem \ref{P:dpcrt} illustrated above is described by the pseudo code of lines 12-23 of Algorithm~\ref{A:dpcrt}.}

@@

\textcolor[rgb]{0,0,1}{\begin{algorithm}[ht!]
	\caption{Data Predictive Control with Regression Trees}
	\label{A:dpcrt}
	\begin{algorithmic}[1]
		\State \textsc{Design Time (Off-Line)}
		\Procedure{Model Training using Dataset Splitting}{}
		\State Set $\X^c$ $\gets$ manipulated features
		\State Set $\X^d$ $\gets$ non-manipulated features
		\State Build $N$ predictive trees with $(\X^d,\Y)$ using Regression Trees algorithm
		\ForAll{trees $\mathcal{T}_j$}
		\ForAll{leaves $\ell_i$ of $\mathcal{T}_j$}
		\State Compute parameters $\beta_{ij}$ in \eqref{E:model_leaf} using convex programming
		\EndFor
		\EndFor
		\EndProcedure
		\State \textsc{Run Time}
		\Procedure{Predictive Control}{}
		\While{$k< k_{\mathrm{stop}}$}
		\ForAll{trees $\mathcal{T}_j$}
		\State Determine the leaf $\ell_i$ using $\X^d$ as in \eqref{E:model_tree}
		\State Obtain the linear model at $\ell_i$ trained in \eqref{E:model_leaf}
		\EndFor
		\State Solve Problem \ref{P:dpcrt} to determine optimal
		\State control actions $u^*_k,\ldots,u^*_{k+j}$
		\State Apply the first input $u(k)=u^*_k$
		\EndWhile
		\EndProcedure
	\end{algorithmic}
\end{algorithm}}

@@

\textcolor[rgb]{0,0,1}{More precisely, in DPC-En we replace each tree in Algorithm~\ref{A:dpcrt} by a forest $\F_j$ of $t$ trees $\T_{1j},\ldots,\T_{tj}$. Each tree $\T_{\kappa j}$ is trained on the basis of a different subset of features $\bar{\X}^d_{\kappa j} \subset \X^d$. As discussed in the previous section, at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state $\left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) \right)$, we can associate the corresponding leaf $\ell_i$ of $\T_{\kappa j}$
\begin{equation}\label{E:model_forest}
	\ell_i = \mathit{f}_{\mathcal{T}_{\kappa j}} \left( d(k+j-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right).
\end{equation}
Also, in each leaf $\ell_i$ of tree $\mathcal{T}_{\kappa j}$ we derive an affine function $\Theta_{i \kappa j}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
\begin{equation}\label{E:model_leaf_forest}
	y(k+j) =  \Theta_{i \kappa j} [1,u(k),\ldots,u(k+j) ]^T.
\end{equation}
}

@@

\textcolor[rgb]{0,0,1}{We can now set up the MPC problem for DPC-En as follows:}
\begin{problem}\label{P:dpcrf}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{j} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
		&                                       & & \epsilon_j \geq  0							                                     \\
		&                                       & & j           =    0,\ldots,N.            									         \\
		\end{aligned}
		\label{E:dpcrf}
	\end{equation}
\end{problem}
\textcolor[rgb]{0,0,1}{Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$. We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$. In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{j} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$. Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{j}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section. The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}

>>>>

\textcolor[rgb]{0,0,1}{The central idea behind DPC is to obtain control-oriented models using machine learning on historical datasets of buildings and formulate the optimal control problem in a way that Receding Horizon Control (RHC) can be solved efficiently. Let an historical dataset $(\X,\Y)$ be given.
We define $\X = \{(x(k),u(k),d(k))\}_{k=1}^n$ the set of predictor variables (or features), i.e. the set of samples $(x(k),u(k),d(k))$ measured a time instants $k=1,\ldots,n$, where $x(k)\in\mathbb{R}^{n_x}$ is the vector of the \emph{measurable} state variables (e.g. the rooms temperatures, power consumption of the building, and others), $u(k)\in\mathbb{R}^{n_u}$ is the vector of the \emph{measurable} input variables (e.g. the set-points and schedule information of the building) and $d(k)\in\mathbb{R}^{n_d}$ is the vector of the \emph{measurable and predictable} disturbance variables (e.g. weather historical data).
We define $\Y = \{y(k)\}_{k=1}^{n}$ the set of response variables, i.e. the set of measured samples $y(k)\in\mathbb{R}^{n_y}$ representing the variables we wish to predict with our model. In the most general case, when we want to predict all measured variables, the response variables are represented by the state variables at the next time step, i.e. $y(k) = x(k+1)$. However, in general, we only wish to predict a subset $y(k) = \bar x(k+1) \subset \mathbb{R}^{n_x}$ of variables, e.g. only room temperatures and power consumption.
Clearly, $|\X| = |\Y| = n$ is the number of time samples of the dataset.\\
We remark that, as explained in Section \ref{secModelingIssuesExample}, the dataset $(\X,\Y)$ does not contain non-measurable states such as temperature of wall layers, ceiling, floor, etc.: nevertheless, our simulated and experimental results show that our data-driven methodology is able to compensate the absence of such variables while providing excellent prediction accuracy.}

@@

\textcolor[rgb]{0,0,1}{Assume that, at time $k$, we want to predict the response variable at time $k+j$, i.e. $y(k+j),\ j\in \{1,\ldots,N\}$: when the data have lots of features interacting in complicated, nonlinear ways, assembling a single global model such as linear or polynomial regression can be difficult, and can lead to poor response predictions.}

@@

\textcolor[rgb]{0,0,1}{Our modification of the Regression Tree algorithm first partitions the features set $\X$ into the sets $\X^c = \{u(k)\}_{k=1}^n\subset\X$, containing the control (or manipulated) variables, and $\X^d = \{(x(k),d(k))\}_{k=1}^n\subset\X$, containing the disturbance and state (or non-manipulated) variables. The  union of the two non-intersecting sets forms the full feature set of training $\X \equiv \X^c \cup \X^d$.
%Our goal is to replace a model-based controller with a data-driven controller, where the latter depends only on the historical sensor data. 
%These measurements could directly represent one or more states in the model-based control framework. We denote these as responses $\Y \in \R$ for training, i.e. a $\Y$ represents a particular response and we can have separate models for multiple responses. We define the number of training samples by $|(\X,\Y)| = n$.
Then the training process is divided into the following two steps, which generate as output a tree $\mathcal{T}_j$ as illustrated in Figure~\ref{F:dpc-sepvars} (left):
\begin{enumerate}
	\item the splitting of the dataset only partitions $\X^d$ (see Appendix 1 and \cite{hastie2009elements} for technical details): this choice is necessary to make our model suitable for control, as we will clarify later on, and also reduces the computational complexity. To each leaf $\ell_i$ will correspond an equivalence class of data samples $\X^d_i$ of the partition of $\X^d$, with $\X^d_i\subset \X^d$: as a consequence at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$
	\begin{equation}\label{E:model_tree}
		\ell_i = \mathit{g}_{\mathcal{T}_j} \left( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right)
	\end{equation}
	if and only if $( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)) \in \X^d_i$.
	\item in each leaf $\ell_i$ of the tree $\mathcal{T}_j$ we derive, solving a convex program over the data samples in the partition $\X^d_i$ associated to leaf $\ell_i$, an affine function, with coefficients $\beta_{ij}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
	\begin{equation}\label{E:model_leaf}
		y(k+j) =  \beta_{ij} [1,u(k),\ldots,u(k+j) ]^T
	\end{equation}
	Clearly the coefficients $\beta_{ij}$ are different for each leaf $\ell_i$ as they are derived from different sets of samples.
\end{enumerate}
As an example, the tree $\mathcal{T}_0$ defines the function $f$ in \eqref{E:GenericModel} to predict $y(k)$ as follows: assuming for simplicity that $\delta_x = \delta_d = 0$, given the measurement of the variables $(x(k), d(k))$ at time $k$, it is possible to determine by \eqref{E:model_tree} the corresponding leaf $\ell_i$ of $\T_0$ and the associated $\beta_{i0}$. The prediction of $y(k)$ is provided by \eqref{E:model_leaf} as an affine function of the control variable $u(k)$.\\
Applying the above procedure for $j=0,\ldots,N$, we build $N$ regression trees $\mathcal{T}_0,\ldots,\mathcal{T}_N$: thus, we have managed to linearize the original model dynamics via black-box modeling.\\ 
Our two-steps training procedure, described by the pseudo code of lines 1-11 of Algorithm~\ref{A:dpcrt}, can be computed off-line: this is an important advantage because the time required to create the model does not affect the control execution in run-time.}

@@

\begin{problem}\label{P:dpcrt}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}     =   \hat \beta_{j} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top                       \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                                   \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										              \\
		&                                       & & \epsilon_j \geq  0							                                                  \\
		&                                       & & j           =    0,\ldots,N,               									                  \\
		\end{aligned}
		\label{E:dpcrt}
	\end{equation}
\end{problem}

\noindent \textcolor[rgb]{0,0,1}{where $\hat \beta_j$ is defined later on. Here, $Q \succeq 0 \in \mathbb{R}^{n_y\times n_y}$ and $R \succeq 0 \in \mathbb{R}^{n_u\times n_u}$ are weight matrices used to trade-off the importance we provide in the minimization to $y$ versus $u$. The slack variables $\epsilon_j$ are added to ensure recursive feasibility: we relax the equality constraint on $y$ allowing violation up to $\epsilon_j$ to guarantee that Problem \ref{P:dpcrt} can provide a solution at each step $k \geq 0$. The weight $\lambda$ is then used to tradeoff the importance we provide in the optimal solution to the constraints violation versus the quadratic term. For example in the following sections we will use $\epsilon_j$ and $\lambda$ to define our tolerance to the violation of prescribed bounds by the rooms' temperature.\\
Clearly different cost functions can be chosen depending upon the application, i.e. they can be linear, nonlinear, etc., obviously changing the complexity of the problem. In the current formulation, the data-driven control problem, is reduced to a convex program which is very easy and efficient to solve in C, C++, Python and Matlab, and thus is easily integrable in SCADA systems.
Indeed, Problem \ref{P:dpcrt} is solved as in the classical MPC formulation, i.e. at each time step $k=1,2,\ldots$ the optimal control sequence $u^*_k,\ldots,u^*_{k+N}$ is computed, and only the first input of the sequence is applied to the system: $u(k) = u^*_k$.\\
Note that each tree $\T_j$ contributes to Problem  \ref{P:dpcrt} with the linear constraint $y_{k+j} = \hat \beta_{j} [1,u^\top_{k},\ldots,u^\top_{k+j} ]^\top$ as a replacement for the state dynamics in the classical MPC formulation. As a consequence, when solving Problem \ref{P:dpcrt} at time $k$, we need to determine the affine functions parameters $\hat \beta_{j}$: it follows by Equation \eqref{E:model_tree} that, to determine $\beta_{ij}, j=0,\ldots,N$ at time $k$, the knowledge of the state and disturbance measurements $\left( d(k-\delta_d),\ldots,d(k+N),x(k-\delta_x),\ldots,x(k) \right)$ is needed. However, the values of $d(k+1),\ldots,d(k+N)$ are not available at time $k$, so we need to use disturbance forecast $\tilde d(k+1),\ldots,\tilde d(k+N)$. Using this information we can narrow down to a leaf in each tree using \eqref{E:model_tree} and thus retrieve the affine model with $\beta_{ij}$ in \eqref{E:model_leaf} for each step $j=0,\ldots,N$ of the horizon, and associate $\hat \beta_j := \beta_{ij}$. The run-time solution of Problem \ref{P:dpcrt} illustrated above is described by the pseudo code of lines 12-23 of Algorithm~\ref{A:dpcrt}.}

@@

\textcolor[rgb]{0,0,1}{\begin{algorithm}[ht!]
	\caption{Data Predictive Control with Regression Trees}
	\label{A:dpcrt}
	\begin{algorithmic}[1]
		\State \textsc{Design Time (Off-Line)}
		\Procedure{Model Training using Dataset Splitting}{}
		\State Set $\X^c$ $\gets$ manipulated features
		\State Set $\X^d$ $\gets$ non-manipulated features
		\State Build $N$ predictive trees with $(\X^d,\Y)$ using Regression Trees algorithm
		\ForAll{trees $\mathcal{T}_j$}
		\ForAll{leaves $\ell_i$ of $\mathcal{T}_j$}
		\State Compute parameters $\hat\beta_j:=\beta_{ij}$ in \eqref{E:model_leaf} using convex programming
		\EndFor
		\EndFor
		\EndProcedure
		\State \textsc{Run Time}
		\Procedure{Predictive Control}{}
		\While{$k< k_{\mathrm{stop}}$}
		\ForAll{trees $\mathcal{T}_j$}
		\State Determine the leaf $\ell_i$ using $\X^d$ as in \eqref{E:model_tree}
		\State Obtain the linear model at $\ell_i$ trained in \eqref{E:model_leaf}
		\EndFor
		\State Solve Problem \ref{P:dpcrt} to determine optimal
		\State control actions $u^*_k,\ldots,u^*_{k+j}$
		\State Apply the first input $u(k)=u^*_k$
		\EndWhile
		\EndProcedure
	\end{algorithmic}
\end{algorithm}}

@@

\textcolor[rgb]{0,0,1}{More precisely, in DPC-En we replace each tree in Algorithm~\ref{A:dpcrt} by a forest $\F_j$ of $t$ trees $\T_{1j},\ldots,\T_{tj}$. Each tree $\T_{\kappa j}$ is trained on the basis of a different subset of features $\bar{\X}^d_{\kappa j} \subset \X^d$. As discussed in the previous section, at any time $k$, given $\delta_d$ autoregressive terms of the disturbances and $\delta_x$ autoregressive terms of the state, we can associate $( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k) )$, to the corresponding leaf $\ell_i$ of $\T_{\kappa j}$
\begin{equation}\label{E:model_forest}
	\ell_i = \mathit{g}_{\mathcal{T}_{\kappa j}} \left( d(k-\delta_d),\ldots,d(k+j),x(k-\delta_x),\ldots,x(k)  \right).
\end{equation}
Also, in each leaf $\ell_i$ of tree $\mathcal{T}_{\kappa j}$ we derive an affine function $\Theta_{i \kappa j}$, that relates the response $y(k+j)$ to the previous control inputs in $\X^c$:
\begin{equation}\label{E:model_leaf_forest}
	y(k+j) =  \Theta_{i \kappa j} [1,u(k),\ldots,u(k+j) ]^T.
\end{equation}
}

@@

\textcolor[rgb]{0,0,1}{We can now set up the MPC problem for DPC-En as follows:}
\begin{problem}\label{P:dpcrf}
	\begin{equation}
		\begin{aligned}
		& \underset{u_{k+j},\epsilon_j}{\mathrm{minimize}} & & \sum_{j=0}^{N} y^\top_{k+j} Q y_{k+j} + u^\top_{k+j} R u_{k+j} + \lambda\epsilon_j \\
		& \mathrm{subject\ to }                 & & y_{k+j}      =  \hat{\Theta}_{j} [1,u_{k},\ldots,u_{k+j} ]^\top                      \\
		&                                       & & u_{k+j}    \in  \mathcal{U}                                                        \\
		&                                       & & |y_{k+j}|  \leq \bar{y}_{k+j} + \epsilon_j 										 \\
		&                                       & & \epsilon_j \geq  0							                                     \\
		&                                       & & j           =    0,\ldots,N,            									         \\
		\end{aligned}
		\label{E:dpcrf}
	\end{equation}
\end{problem}
\noindent \textcolor[rgb]{0,0,1}{where $\hat{\Theta}_{j}$ is defined later on. Note that, as in the DPC-RT, the crucial problem when solving Problem \ref{P:dpcrf} is determining all parameters $\hat{\Theta}_{j}$ at time $k$. We first proceed similarly to DPC-RT by using forecast of disturbances and the state measurements to determine for each tree $\T_{\kappa j}$ the corresponding leaf $\ell_i$ and thus the affine model $\Theta_{i \kappa j}$, with $\kappa = 1,\ldots,t$ and $j = 0,\ldots,N$. In DPC-En, however, we have $t$ affine models for each prediction step $j$, thus we define $\hat{\Theta}_{j} = \frac{1}{t}\sum\limits_{\kappa = 1}^{t} \Theta_{i \kappa j}$ averaging the affine models associated to all trees of forest $\F_j$. Once we are able to determine at each time $k$ the parameters $\hat{\Theta}_{j}$ of the constraints of Problem \ref{P:dpcrf}, the optimal input $u^*_k$ can be computed as described in the previous section. The overall procedure is sketched in Figure \ref{F:dpc-algo-rf}.}
